---
title: "Stratford"
output:
  md_document:
    variant: markdown_github
---

I started going to the [Stratford Festival](https://www.stratfordfestival.ca/) with my grandparents when I turned seven. Honestly, I hated that first time. Shakespeare and seven years old go together about as well as peanut butter and snot. But it meant a lot to my grandparents. So, I signed up to do it again when I was eight. 

I hated that second time, too. But the third time we did the Festival I hated it a little bit less. And the fourth time I actually kind of enjoyed it. 

By the time I was in grade 8 Shakespeare had become a significant enough part of my life that I decided to write my Speech on *The Bard*. I can't remeber it exactly but it was something like, "Why you should give Shakespeare a chance in High School" (I was really cool in grade school...)

Now, Shakespeare at the Stratford Festival has become an incredibly rich tradition for me and my grandparents. We've done the festival 16-years in a row! And we're just now starting to plan our seventeenth adventure. The only problem: We're running out of plays to see! And it's getting harder and harder to remember what we've actually seen.

This year, Stratford is running *Twelfth Night*, *Romeo and Juliet* and *Timon of Athens*. I was sure we had seen  *Twelfth Night* and *Romeo and Juliet* so our decision was pretty easy. But still, I was kind of curious about what was still left... 

```{r, warning = FALSE, message = FALSE, include = FALSE}
# load packages
library(tidyverse)
library(rvest)
library(purrr)
library(stringr)
library(knitr)
library(forcats)

# knitr options
opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r}
# base url
URL <- read_html("https://en.wikipedia.org/wiki/Stratford_Shakespeare_Festival_production_history")

# selectors
COMEDY_SELECTOR <- "ul:nth-child(144) li"
HISTORY_SELECTOR <- "ul:nth-child(146) li"
TRAGEDY_SELECTOR <- "ul:nth-child(148) li"
```

```{r}
# play scrape and cleaner
get_plays <- function(HTML_SELECTOR) {

    df <- URL %>% 
        html_nodes(HTML_SELECTOR) %>% 
        html_text() %>% 
        as_tibble() %>% 
        separate(value, into = c("play", "year"), sep = "\\s\\-\\s") %>% 
        separate(year, into = paste(1:25), sep = ",", fill = "right") %>% 
        gather(edition, year, -play) %>% 
        drop_na(year) %>% 
        arrange(play) %>% 
        group_by(play) %>% 
        mutate(edition = as.integer(edition), year = as.integer(year)) %>% 
        mutate(gap = year - lag(year))
    
    return(df)
}

# get plays
comedies <- get_plays(COMEDY_SELECTOR) %>% mutate(type = "Comedy")
histories <- get_plays(HISTORY_SELECTOR) %>% mutate(type = "History")
tragedies <- get_plays(TRAGEDY_SELECTOR) %>% mutate(type = "Tragedy")
```

```{r}
# full list
shakespeare <- comedies %>% 
    bind_rows(histories) %>% 
    bind_rows(tragedies)
```

## Data Investigation

Simply pulling and graphing the runs for every play for single season didn't really tell me much...

```{r shakespeare_01}
# attempt 1
shakespeare %>% 
    ggplot(aes(y = play, x = year)) + 
    geom_point()
```

Trying to get fancy with lines instead of points didn't help either...

```{r shakespeare_02}
# attempt 2
shakespeare %>% 
    ggplot(aes(y = play, x = year)) + 
    geom_point(shape = 124, size = 3)
```

Zooming in the *Histories* was interesting and telling. Just look at how the *Henry* plays are bunched up and played at the same time during the same season...

```{r shakespeare_03}
# attempt 3
shakespeare %>% 
    filter(type == "History") %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = play, x = year)) + 
    geom_point(shape = 124, size = 10)
```

But zooming back out and colouring all the genres on the same vertical line plot really didn't work...

```{r shakespeare_04}
# attempt 4
shakespeare %>% 
    # filter(type == "History") %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = play, x = year, color = type)) + 
    geom_point(shape = 124, size = 10)
```

Faceting the genres with three columns was a disaster...

```{r shakespeare_05}
# attempt 5
shakespeare %>% 
    # filter(type == "History") %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = play, x = year, color = type)) + 
    geom_point(shape = 124, size = 10) +
    facet_wrap(~type, scales = "free_y")
```

And a one-column facet wasn't much better...

```{r shakespeare_06}
# attempt 6
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = play, x = year)) + 
    geom_point(shape = 124, size = 10) +
    facet_wrap(~type, scales = "free_y", ncol = 1)
```

I was starting to get somewhere when I reverted back to points (instead of lines) and when I reordered the data accoring to the most recent runs of each play...

```{r shakespeare_07}
# attempt 7
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, fill = type)) + 
    geom_point(shape = 21, size = 2)
```

Moving the y-axis text over...

```{r shakespeare_08}
# attempt 8
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, fill = type)) + 
    geom_point(shape = 21, size = 2) + 
    scale_y_discrete(position = "right")
```

And trimming the data to 2000 helped immensely...

```{r shakespeare_09}
# attempt 9
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, fill = type)) + 
    geom_point(shape = 21, size = 2) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(limits = c(2000, 2017)) + 
    theme_minimal() +
    theme(legend.position = "none") + 
    labs(x = "", y = "")
```

At this point, I was able to reverse out exactly the play that me and my grandparents had seen and when.

```{r shakespeare_10}
# attempt 10
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, fill = type)) + 
    geom_point(shape = 21, size = 2) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(
        limits = c(2000, 2018), 
        breaks = seq(2000, 2018, 2),
        labels = c("2000", paste0("'0", seq(2, 8, 2)), paste0("'", seq(10, 18, 2)))) + 
    theme_minimal() +
    theme(legend.position = "none") + 
    labs(title = "Shakespeare Playbill at Stratford", x = "", y = "")
```

After trying to clean up the styling...

```{r shakespeare_11}
# attempt 11
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, fill = type)) + 
    geom_point(shape = 21, size = 2) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(
        limits = c(2000, 2018), 
        breaks = seq(2000, 2018, 2),
        labels = c("2000", paste0("'0", seq(2, 8, 2)), paste0("'", seq(10, 18, 2)))) + 
    labs(title = "Shakespeare Playbill at Stratford", x = "", y = "") + 
    theme(
        plot.title = element_text(colour = "#3C3C3C", size = 12),
        axis.text = element_text(colour = "#3C3C3C", size = 10),
        panel.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        plot.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
```

And trying to mess with the colors (gross)...

```{r shakespeare_12}
# attempt 12
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, color = type)) + 
    geom_point(shape = 16, size = 3, color = "#3C3C3C") + 
    geom_point(shape = 16, size = 2) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(
        limits = c(2000, 2018), 
        breaks = seq(2000, 2018, 2),
        labels = c("2000", paste0("'0", seq(2, 8, 2)), paste0("'", seq(10, 18, 2)))) + 
    scale_color_manual(values = c("History" = "yellow", "Comedy" = "green", "Tragedy" = "red")) + 
    labs(title = "Shakespeare at The Stratford Festival", x = "", y = "") + 
    theme(
        plot.title = element_text(colour = "#3C3C3C", size = 12),
        axis.text = element_text(colour = "#3C3C3C", size = 10),
        panel.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        plot.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
```

I finally...

```{r shakespeare_13}
# attempt 13
shakespeare %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, color = type)) + 
    geom_point(shape = 16, size = 3, color = "#3C3C3C", alpha = 0.75) + 
    geom_point(shape = 16, size = 2) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(
        limits = c(2000, 2018), 
        breaks = seq(2000, 2018, 2),
        labels = c("2000", paste0("'0", seq(2, 8, 2)), paste0("'", seq(10, 18, 2)))) + 
    scale_color_manual(values = c("History" = "#39a9db", "Comedy" = "#f39237", "Tragedy" = "#d63230")) + 
    labs(title = "Shakespeare at The Stratford Festival", x = "", y = "") + 
    theme(
        plot.title = element_text(colour = "#3C3C3C", size = 12),
        axis.text = element_text(colour = "#3C3C3C", size = 10),
        panel.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        plot.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "none")
```

```{r}
# plays we've seen
humber <- tribble(
    ~play,~edition,~year,~gap,~type,
    "A Midsummer Night's Dream",10,2004,5,"Comedy",
    "All's Well That Ends Well",5,2002,14,"Comedy",
    "As You Like It",10,2005,5,"Comedy",
    "Love's Labour's Lost",8,2003,11,"Comedy",
    "The Comedy of Errors",7,2007,12,"Comedy",
    "The Merchant of Venice",7,2001,5,"Comedy",
    "The Taming of the Shrew",10,2015,7,"Comedy",
    "The Tempest",7,2010,5,"Comedy",
    "Twelfth Night",10,2006,5,"Comedy",
    "Henry IV, Part 2",7,2016,10,"History",
    "Richard III",7,2011,9,"History",
    "Cymbeline",4,2012,8,"Tragedy",
    "Hamlet",8,2008,8,"Tragedy",
    "King Lear",10,2014,7,"Tragedy",
    "Macbeth",10,2009,5,"Tragedy",
    "Romeo and Juliet",10,2013,5,"Tragedy",
    "Timon of Athens",4,2017,13,"Tragedy") %>% 
    mutate(seen = 1) %>% 
    select(play, year, seen)

# complete join
shakey <- shakespeare %>% 
    left_join(humber, by = c("play", "year")) %>% 
    mutate(seen = ifelse(!is.na(seen), 1, 0))

```

...arrived at this polished and finished version. 

Here I present every play that me and my grandparents have seen so far (the Xs)...

```{r shakespeare_14}
# attempt 14
shakey %>% 
    group_by(play) %>% 
    mutate(latest = max(year)) %>% 
    ggplot(aes(y = fct_reorder(play, latest), x = year, color = type, shape = factor(seen))) + 
    geom_point(size = 3, color = "#3C3C3C", alpha = 0.75) + 
    geom_point(size = 2) + 
    scale_shape_manual(values = c(16, 4)) + 
    scale_y_discrete(position = "right") + 
    scale_x_continuous(
        limits = c(2000, 2018), 
        breaks = seq(2000, 2018, 2),
        labels = c("2000", paste0("'0", seq(2, 8, 2)), paste0("'", seq(10, 18, 2)))) + 
    scale_color_manual(values = c("History" = "#39a9db", "Comedy" = "#f39237", "Tragedy" = "#d63230")) + 
    labs(title = "Shakespeare at The Stratford Festival", x = "", y = "") + 
    theme(
        plot.title = element_text(colour = "#3C3C3C", size = 10),
        axis.text = element_text(colour = "#3C3C3C", size = 8),
        panel.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        plot.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.grid.major = element_line(linetype = 1, size = 0.25, color = "#D0D0D0"),
        panel.grid.minor = element_line(linetype = 1, size = 0.25, color = "#D0D0D0"),
        axis.ticks = element_blank(),
        legend.position = "none")

```

The cool thing about this graph is you can infer what Stratford might play in the seasons to come (*Coriolanus* and *The Two Noble Kinsmen* is due for a run)...

If you're interested in seeing the code for each graph and how I pulled the data from Wikipedia it's available on [GitHub](https://github.com/maxhumber/maxhumber.com/tree/master/_R)
