---
title: "The Goodreads API"
output:
  md_document:
    variant: markdown_github
---

It's December 23rd and I've only read 49 books. Shit. There's still time, but it's definitely getting dicey. I'm about halfway through three books right now so I think I'll be able to pull it off. Fingers crossed.

Of course, last year I did 52 books in 52 weeks and remember sitting pretty just before Christmas. 

As I've been logging all my activity on [Goodreads](https://www.goodreads.com/user/show/16626766-max) I thought it would be neat to plug into the API and compare my reading progress between the years. To see if I read at the same pace or if there's some sort of seasonality in my reading habits.

If you happen to use Goodreads and want to do the same here's how:

# Setup

(Man, I love Hadley)

```{r, warning = FALSE, message = FALSE}
# load packages
library(httr)
library(tidyverse)
library(stringr)
library(xml2)
library(viridis)
library(knitr)

# knitr options
opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```

# API Guide

```{r, include = FALSE}
# censored
API_KEY <- "K97BcQqbrP9X8e2gY4Z2g"
GR_ID <- "16626766"
```

I've censored my `API_KEY` and `GR_ID` but if you replace the `"XXXXXXXXXXXXX"`s with your KEY and your ID you should be good to go!

```{r}
# API_KEY <- "XXXXXXXXXXXXX"
# GR_ID   <- "XXXXXXXXXXXXX"

# https://www.goodreads.com/review/list?key=K97BcQqbrP9X8e2gY4Z2g&id=16626766&v=2&shelf=read&per_page=200
URL <- "https://www.goodreads.com/review/list?"

# v: 2
# id: Goodreads id of the user
# shelf: read, currently-reading, to-read, etc. (optional)
# sort: title, author, cover, rating, year_pub, date_pub, date_pub_edition, date_started, date_read, date_updated, date_added, recommender, avg_rating, num_ratings, review, read_count, votes, random, comments, notes, isbn, isbn13, asin, num_pages, format, position, shelves, owned, date_purchased, purchase_location, condition (optional)
# search[query]: query text to match against member's books (optional)
# order: a, d (optional)
# page: 1-N (optional)
# per_page: 1-200 (optional)
# key: Developer key (required).
```

# Get Shelf

This is where the heavy lifting `GETS` done. I'm leaning on `httr` and `XML2` to parse the API responses.

```{r}
get_shelf <- function(GR_ID) {
    shelf <- GET(URL, query = list(v = 2, key = API_KEY, id = GR_ID, shelf = "read", per_page = 200))
    shelf_contents <- content(shelf, as = "parsed")
    return(shelf_contents)
}

shelf <- get_shelf(GR_ID)

get_df <- function(shelf) {

    title <- shelf %>% 
        xml_find_all("//title") %>% 
        xml_text()
    
    rating <- shelf %>% 
        xml_find_all("//rating") %>% 
        xml_text()
    
    added <- shelf %>% 
        xml_find_all("//date_added") %>% 
        xml_text()
    
    started <- shelf %>% 
        xml_find_all("//started_at") %>% 
        xml_text()
    
    read <- shelf %>% 
        xml_find_all("//read_at") %>% 
        xml_text()
    
    df <- tibble(
        title, rating, added, started, read)
    
    return(df)
}

df <- get_df(shelf)
```

# Clean

After getting the XML data into my IDE I tabled and cleaned the data with `dplyr` and `tidyr`.

```{r}
get_books <- function(df) {

    books <- df %>% 
        gather(date_type, date, -title, -rating) %>% 
        separate(date, 
            into = c("weekday", "month", "day", "time", "zone", "year"), sep = "\\s", fill = "right") %>% 
        mutate(date = str_c(year, "-", month, "-", day)) %>% 
        select(title, rating, date_type, date) %>% 
        mutate(date = as.Date(date, format = "%Y-%b-%d")) %>% 
        spread(date_type, date) %>% 
        mutate(title = str_replace(title, "\\:.*$|\\(.*$|\\-.*$", "")) %>% 
        mutate(started = ifelse(is.na(started), as.character(added), as.character(started))) %>% 
        mutate(started = as.Date(started)) %>% 
        mutate(rating = as.integer(rating))
    
    return(books)
}

books <- get_books(df)
```

```{r, include = FALSE}
book_idx <- books %>% 
    mutate(read_idx = 0) %>% 
    mutate(started_idx = as.numeric(read - started) + 1) %>% 
    mutate(added_idx = as.numeric(read - added) + 1) %>% 
    mutate(started_idx = ifelse(started_idx <= 0, 1, started_idx)) %>% 
    mutate(added_idx = ifelse(added_idx <= 0, 1, added_idx)) %>% 
    arrange(read) %>% 
    mutate(id = row_number(read)) %>% 
    complete(id = seq(1, 100, 1)) %>% 
    mutate(section = cut(id, breaks = seq(0, 500, 25)))
```

# Compare

All of that get to this graph:

```{r, echo = FALSE}
books %>% 
    mutate(year = format(read, "%Y")) %>% 
    arrange(year, read) %>% 
    mutate(one = 1) %>% 
    group_by(year) %>% 
    mutate(cum_sum = cumsum(one)) %>% 
    mutate(read_week = as.numeric(format(read, "%U")) + 1) %>%
    ggplot(aes(x = read_week, y = cum_sum, color = year, group = year)) + 
    geom_abline(intercept = 0, slope = 1, 
        linetype = 3, colour = "#3C3C3C", size = 0.25) + 
    geom_line(size = 1) + 
    scale_y_continuous(
        position = "right",
        name = NULL,
        limits = c(0, 52),
        breaks = c(0, 13, 26, 39, 52),
        labels = c(
            "0", "13 Books\n(25%)", "26 Books\n(50%)", 
            "39 Books\n(75%)", "52 Books\n(100%)")) + 
    scale_color_manual(values = c("#999999", "red")) + 
    scale_x_continuous(
        name = NULL,
        limits = c(1, 52),
        breaks = c(1, 13, 26, 39, 52),
        labels = c("Q1", "Q2", "Q3", "Q4", "EOY")) + 
    theme(
        legend.title = element_blank(),
        legend.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        plot.background = element_rect(fill = "#F0F0F0", colour = "#F0F0F0"),
        panel.grid.major = element_line(colour = "#D0D0D0", size = .25, linetype = 1), 
        axis.ticks = element_blank(),
        legend.position = "top", 
        axis.text = element_text(size = 11, colour = "#535353"),
        panel.grid.minor = element_blank())
```

It's funny to see that I started strong in both years and fell off in March. Though I recovered somewhat in 2015, Summer 2016 was a bad season for reading, apparently. 

Looks like I was finished 52 books by December 21 last year. Whoops. Oh well, I still think I can mad rush it to the finish line. 
